### 2012-07-25 Doc/VRIPHYS2012/vriphys2012.tex
finishing section 2
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@432 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\noindent
where the points $P_i$ are named as in Figure \ref{fig-bezier}.

% TODO: explain the equation eq-central? would require an image

@@

With 10 control points and 3 DOFs per point it is a total of 30 DOFs for triangle.
This puts ... blah blah.
We propose a reduction of the number of degrees of freedom by using 3 nodes with 6 degrees of freedom
by creating a kinematic link between the control points.
Each of the edge control points is attached to nearest corner node. See Figure
\ref{fig-segments} for the correspondence between edge points and
associated corner nodes. 
In the following text the index $j \in [4;10]$ refers to edge points and index $i \in [1;3]$ to the associated corner point/node.

@@

For the central point we again apply interpolation of \eqref{eq-central}.

@@

\subsection{Deflection function} % {{{

% TODO: we need to justify how we decide to move the control points

\CD{It's more about derivatives of the kinematic relation, right ?}

We recall the equation for central node \eqref{eq-central} and
considering how are the edge points of control mesh connected to corner points with
segments $P_{i-j}$:

\begin{equation}
    P_{i-j} = P_j - P_i, \quad j \in [4;9], i \in [1;3].
\end{equation}

We now add small variation to the position ($U_1, U_2, U_3$) and orientation ($\theta_1, \theta_2, \theta_3$) of corner nodes to the equation of the element \eqref{eq-cubicbez}:

\begin{equation}
    \begin{split}
        & n_1 (P_1 + U_1)
        + n_2 (P_2 + U_2)
        + n_3 (P_3 + U_3) \\
        & + n_4 (P_1 + U_1 + P_{1-4} + P_{1-4} \times \theta_1)
        + n_5 (P_1 + U_1 + P_{1-5} + P_{1-5} \times \theta_1) \\
        & + n_6 (P_2 + U_2 + P_{2-6} + P_{2-6} \times \theta_2)
        + n_7(P_2 + U_2 + P_{2-7} + P_{2-7} \times \theta_2) \\
        & + n_8 (P_3 + U_3 + P_{3-8} + P_{3-8} \times \theta_3)
        + n_9 (P_3 + U_3 + P_{3-9} + P_{3-9} \times \theta_3) \\
        & + n_{10} \frac{1}{3} \left(
        \sum_{j=4}^9 (P_i + U_i + P_{i-j} + P_{i-j} \times \theta_i) - \sum_{i=1}^3 (P_i + U_i)
        \right) \\
%
        & = 
    \sum_{i=1}^{10} n_i P_i
      + n_1 U_1
      + n_2 U_2
      + n_3 U_3 \\
      & + n_4 (U_1 + P_{1-4} \times \theta_1)
      + n_5 (U_1 + P_{1-5} \times \theta_1)
      + n_6 (U_2 + P_{2-6} \times \theta_2) \\
      & + n_7 (U_2 + P_{2-7} \times \theta_2)
      + n_8 (U_3 + P_{3-8} \times \theta_3)
      + n_9 (U_3 + P_{3-9} \times \theta_3) \\
      & + n_{10} \frac{1}{3} (U_1 + U_2 + U_3 + \sum_{j=4}^9 P_{i-j} \times \theta_i) \\
%
      & = T + U_1 (n_1 + n_4 + n_5 + \frac{1}{3} n_{10}) \\
      & + U_2 (n_2 + n_6 + n_7 + \frac{1}{3} n_{10}) \\
      & + U_3 (n_3 + n_8 + n_9 + \frac{1}{3} n_{10}) \\
      & + \sum_{j=4}^9 n_j P_{i-j} \times \theta_i
      + n_{10} \frac{1}{3} \sum_{j=4}^9 P_{i-j} \times \theta_i \\
%
      & = T + u
  \end{split}
\end{equation}

\CD{I would use the following notation: small variation of the node $j$ position:   $\delta \mathbf{x}_j$ for the translation and $\delta \boldsymbol{\theta}_j$ for the rotation.
This small variation of the position creates a small variation of the control points $i$ position $\delta P_i$ that are on the edges:
$$
\delta P_i =  \delta \mathbf{x}_j +  P_{i-j} \times \delta \boldsymbol{\theta}_j
$$
 Variation of node $10$ will be:
$$
\delta P_{10} =  \sum_{j=1}^3  \frac{1}{3}(\delta \mathbf{x}_j +  P_{10-i} \times \delta \boldsymbol{\theta}_i)
$$
%
Thus, finally, it will creates a variation of the point position on the surface $\delta T$...
$$
\delta T =   \sum_{i=1}^{10} n_i  \delta P_i
$$
which makes:
\begin{equation}
    \begin{split}
\delta T =  (n_1 + n_4 + n_5 + \frac{1}{3} n_{10})  \delta \mathbf{x}_1 \\
 + (n_2 + n_6 + n_7 + \frac{1}{3} n_{10})  \delta \mathbf{x}_2 \\
 + (n_3 + n_8 + n_9 + \frac{1}{3} n_{10})  \delta \mathbf{x}_3 \\
+  \sum_{i=4}^9  \sum_{j=1}^3 n_i P_{i-j} \times \delta \theta_j \\
 +  n_{10} \frac{1}{3} \sum_{j=1}^3 P_{10-j} \times \theta_j
  \end{split}
\end{equation}
%
}

\noindent
thus computing the function of deflection $u$ 
\CD{ or $\delta T$ }.

We have simplified the notation by leaving the variable $i$ free in the above sums involving $P_{i-j}$.
Note however that the expressions are sound and unambiguous because there is only one possible $i$ for each $j$ depending on the connection of segments to the corner nodes (see fig. \ref{fig-segments}).

@@

...

Using the plate theory we can now compute the strain-displacement matrices
from the displacement field. From the Cauchy's strain tensor for the
elastic membrane:

\begin{equation}\label{eq-cauchy}
    J_m = \left[ \begin{matrix}
        \deriv{u_x}{x} \\
        \deriv{u_y}{y} \\
        \deriv{u_x}{y} + \deriv{u_y}{x}
    \end{matrix} \right]
\end{equation}

\noindent
and from Kirchhoff-Love theory for thin plates:

>>>>

\noindent
where the points $P_i$ are named as in Figure \ref{fig-bezier}.
% Equation eq-central is better explained below

@@

With 10 control points and 3 DOFs per point it is a total of 30 DOFs for triangle.
Simulation of such complex system would significantly reduce the sparsity of stiffness matrix.
We propose a reduction of the number of degrees of freedom by using 3 nodes with 6 degrees of freedom
by creating a kinematic link between the control points.
Each of the edge control points is attached to nearest corner node. See Figure
\ref{fig-segments} for the correspondence between edge points and
associated corner nodes. 
In the following text the index $j \in [4;10]$ refers to edge points and index $i \in [1;3]$ to the associated corner point/node (unless $i$ is explicitly used differently).

@@

For the central point we again apply interpolation of \eqref{eq-central} which is kind of skinning function:

\begin{equation}\label{eq-central}
    \begin{split}
        P_{10} = & \frac{1}{3}(\sum_{i=4}^9 P_i - \sum_{i=1}^3 P_i)\\
               = & \frac{1}{3} \left[ P_1 + (P_{1-4} + P_{1-5}) \right] + \\
                & + \frac{1}{3} \left[ P_2 + (P_{2-6} + P_{2-7}) \right] + \\
                & + \frac{1}{3} \left[ P_3 + (P_{3-8} + P_{3-9}) \right]
    \end{split}
\end{equation}

@@

\subsection{Deflection Function} % {{{

% explicit {{{
%We recall the equation for central node \eqref{eq-central} and
%considering how are the edge points of control mesh connected to corner points with
%segments $P_{i-j}$:
%
%\begin{equation}
%    P_{i-j} = P_j - P_i, \quad j \in [4;9], i \in [1;3].
%\end{equation}
%
%We now add small variation to the position ($U_1, U_2, U_3$) and orientation ($\theta_1, \theta_2, \theta_3$) of corner nodes to the equation of the element \eqref{eq-cubicbez}:
%
%\begin{equation}
%    \begin{split}
%        & n_1 (P_1 + U_1)
%        + n_2 (P_2 + U_2)
%        + n_3 (P_3 + U_3) \\
%        & + n_4 (P_1 + U_1 + P_{1-4} + P_{1-4} \times \theta_1)
%        + n_5 (P_1 + U_1 + P_{1-5} + P_{1-5} \times \theta_1) \\
%        & + n_6 (P_2 + U_2 + P_{2-6} + P_{2-6} \times \theta_2)
%        + n_7(P_2 + U_2 + P_{2-7} + P_{2-7} \times \theta_2) \\
%        & + n_8 (P_3 + U_3 + P_{3-8} + P_{3-8} \times \theta_3)
%        + n_9 (P_3 + U_3 + P_{3-9} + P_{3-9} \times \theta_3) \\
%        & + n_{10} \frac{1}{3} \left(
%        \sum_{j=4}^9 (P_i + U_i + P_{i-j} + P_{i-j} \times \theta_i) - \sum_{i=1}^3 (P_i + U_i)
%        \right) \\
%%
%        & = 
%    \sum_{i=1}^{10} n_i P_i
%      + n_1 U_1
%      + n_2 U_2
%      + n_3 U_3 \\
%      & + n_4 (U_1 + P_{1-4} \times \theta_1)
%      + n_5 (U_1 + P_{1-5} \times \theta_1)
%      + n_6 (U_2 + P_{2-6} \times \theta_2) \\
%      & + n_7 (U_2 + P_{2-7} \times \theta_2)
%      + n_8 (U_3 + P_{3-8} \times \theta_3)
%      + n_9 (U_3 + P_{3-9} \times \theta_3) \\
%      & + n_{10} \frac{1}{3} (U_1 + U_2 + U_3 + \sum_{j=4}^9 P_{i-j} \times \theta_i) \\
%%
%      & = T + U_1 (n_1 + n_4 + n_5 + \frac{1}{3} n_{10}) \\
%      & + U_2 (n_2 + n_6 + n_7 + \frac{1}{3} n_{10}) \\
%      & + U_3 (n_3 + n_8 + n_9 + \frac{1}{3} n_{10}) \\
%      & + \sum_{j=4}^9 n_j P_{i-j} \times \theta_i
%      + n_{10} \frac{1}{3} \sum_{j=4}^9 P_{i-j} \times \theta_i \\
%%
%      & = T + u
%  \end{split}
%\end{equation}
%}}}

A small variation $\delta \mathbf{x}_i$ of the node $i$ position and $\delta \mathbf{\theta}_i$ for the rotation produces the following small variation of position corner points $\delta P_i$ and edge points $\delta P_j$:

\begin{eqnarray}
    \delta P_i & = & \delta \mathbf{x}_i \\
    \delta P_j & = & \delta \mathbf{x}_i +  P_{i-j} \times \delta \mathbf{\theta}_i
\end{eqnarray}

\noindent
and variation of central point $\delta P_{10}$:

\begin{equation}
    \delta P_{10} = \frac{1}{3} (\sum_{i=1}^3 \delta \mathbf{x}_i + \sum_{j=4}^9 P_{i-j} \times \delta \mathbf{\theta}_i)
\end{equation}

Thus, finally, it creates a variation of the point position on the surface $\delta T$:
\begin{equation}
    \delta T = \sum_{i=1}^{10} n_i  \delta P_i
\end{equation}

\noindent
or:

\begin{equation}
    \begin{split}
        \delta T = &   (n_1 + n_4 + n_5 + \frac{1}{3} n_{10})  \delta \mathbf{x}_1 + \\
                   & + (n_2 + n_6 + n_7 + \frac{1}{3} n_{10})  \delta \mathbf{x}_2 + \\
                   & + (n_3 + n_8 + n_9 + \frac{1}{3} n_{10})  \delta \mathbf{x}_3 + \\
                   & + \sum_{j=4}^9 P_{i-j} \times \delta \mathbf{\theta}_j + \\
                   & + n_{10} \frac{1}{3} \sum_{j=4}^9 P_{i-j} \times \mathbf{\theta}_i
    \end{split}
\end{equation}

\noindent
thus computing the physical deflection $u$ as variation of the point position $\delta T = u$.

We have simplified the notation by leaving the variable $i$ free in the above sums ($\sum_{j=4}^9$) involving $P_{i-j}$.
Note however that the expressions are sound and unambiguous because there is only one possible $i$ for each $j$ depending on the connection of segments to the corner nodes (see fig. \ref{fig-segments}).

@@

Using the plate theory we now can compute the strain-displacement matrices
from the displacement field $u$. The Cauchy's strain tensor for the
elastic membrane is defined as:

\begin{equation}\label{eq-cauchy}
    J_m = \left[ \begin{matrix}
        \deriv{u_x}{x} \\
        \deriv{u_y}{y} \\
        \deriv{u_x}{y} + \deriv{u_y}{x}
    \end{matrix} \right]
\end{equation}
.

We apply Kirchhoff-Love theory for thin plates to obtain strain-displacement matrix for bending plate element:
